Transforming Optimal Tetrahelices between the Boerdijk-Coxeter Helix and a Planar-faced Tetrahelix

### 2018-04-16 rfinaldoc/JMR-17-1184-paper.tex
removed highlights

% \documentclass[review]{siamonline1116}
\documentclass[twocolumn,10pt]{asme2ej}
% \usepackage{geometry} % see geometry.pdf on how to lay out the page. There's lots.
% \usepackage{hyperref}
\usepackage{graphicx}
\usepackage{gensymb}
% \usepackage[affil-it]{authblk}
% \usepackage[toc,page]{appendix}
% \usepackage{pifont}
\usepackage{amsmath}
% \usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{hyperref}
\usepackage{cleveref}

@@

\newcommand{\highlighttext}[1] {\textcolor{red}{#1}}

@@

\title{\highlighttext{Transforming Optimal Tetrahelices} between the Boerdijk-Coxeter Helix and a Planar-faced Tetrahelix}
\author{Robert L. Read
  \affiliation{
    Founder, Public Invention \\
    Austin, TX, 78704 \\
    Email: \href{mailto:read.robert@gmail.com}{read.robert@gmail.com} 
    }
}

@@

%% TODO: Add something about Tetrobots into abstract.
\begin{abstract}
  The Boerdijk-Coxeter helix (BC helix, or tetrahelix) is a
  face-to-face stack of regular tetrahedra forming a helical column.  Treating the edges of
  these tetrahedra as structural members creates an attractive and
  inherently rigid space frame, and therefore is interesting to architects,
  mechanical engineers, and roboticists.  A formula is developed that matches the
  visually apparent helices forming the outer rails of the BC helix.
  This formula is generalized to a formula convenient to designers.
  Formulae for 
  computing the
  parameters that give proven edge-length minimax-optimal tetrahelices
  are given,
  \highlighttext{
  allowing transformation through a continuum of 
  of optimum tetrahelices of varying curvature while maximizing regularity.}
  The endpoints of this continuum are the BC helix and
  a structure of zero curvature, the \emph{equitetrabeam}.
 Only one out of three members in the system change their length \highlighttext{ to transform
  the structure into }
  any point in the continuum.
  Numerically finding the rail angle from the equation for
  pitch allows optimal tetrahelices of any pitch to be designed. 
  An interactive tool for such design and experimentation is provided: \url{https://pubinv.github.io/tetrahelix/}.
  A formula for the inradius of optimal tetrahelices is given.
  The continuum allows a regular Tetrobot supporting a length change of less than $16\%$ in the
  BC configuration to untwist into a hexapodal or $n$-podal robot
  to use standard gaits.
\end{abstract}

@@

The Boerdijk-Coxeter helix\cite{coxeter1985simplicial} (BC helix) (see \cref{fig:closeup,fig:closeuporthog}), is
a face-to-face stack of tetrahedra that winds about a straight axis.
Architects, structural engineers, and roboticists are inspired
by and follow such regular mathematical models.
However, since they can also build structures and
machines from members of differing or even dynamically changing length, it is
useful to develop the mathematics of structures formed from tetrahedra
where \highlighttext{one relaxes} regularity.

@@

Buckminster Fuller called the BC helix a \emph{tetrahelix}\cite{fuller1982synergetics},
a term now commonly used. \highlighttext{This paper} reserves \emph{BC helix} to mean the purely
regular structure and use \emph{tetrahelix} to refer
to any structure isomorphic to the BC helix.

@@

A BC helix does not rest stably on a plane. It is convenient to
be able to ``untwist'' it and to form a tetrahelix space frame that has a
flat planar surface. By making length changes in a certain way, a \highlighttext{tetrahelix untwists} to form a \emph{tetrabeam} which has planar faces
and has, for example, an equilateral triangular profile. This paper
develops the equations needed to untwist the tetrahelix. All math
developed here is available in JavaScript and demonstrated by an interactive
design website \url{https://pubinv.github.io/tetrahelix/}\cite{readtetrahelix}
\highlighttext{and a video\cite{tetrahelixvideo}},
from which \cref{fig:closeup,fig:closeuporthog} and later figures in this paper 
are taken.

@@

\Cref{fig:series} displays a continuum of tetrahelices optimal \highlighttext{in the sense of being
as regular as possible},
which is the main result of this
paper. The closest helix is the BC helix, and the furthest
is the equitetrabeam, defined in \cref{sec:equitetrabeam} and \cref{fig:naming,fig:equitetrabeam}.

@@

However, a robot must move, and so it is interesting to consider the
transmutations of these regular geometries. Developing a functional
gait for our physical Tetrobot (the 7-Tet Glussbot) was in fact the motivation
for this work.
By changing only the length of only one third of the members,
\highlighttext{one} can dynamically untwist a physical Tetrobot
forming the Boerdijk-Coxeter configuration into the equitetrabeam which rests flat on the plane,
as shown in figs. \highlighttext{4 and} 5.

By untwisting the tetrahelix so that is has a planar surface resting on the ground,
each vertex touching the ground \highlighttext{becomes} a foot or pseudopod. A robot can thus
become a hexapod or $n$-pod Tetrobot, and the already well-developed approaches to
hexapod gaits may be applied to make the robot walk or crawl.

@@

\highlighttext{
The remainder of this  paper:
\begin{enumerate}
\item provides a new formulation for the BC helix that is more natural for engineers,
\item defines a concept of minimum maximum edge-length optimality that maintains
  the structure as close to regularity as possible,
  \item proves that all
    optimal tetrahelices are in fact triple helices,
\item provides formulae for constructing optimal tetrahelices of different twists or for optimally
    transforming a
    robot from a twisted to a planar configuration in a way that maintains optimum regularity,
\item gives a formula for the inradius of an optimal tetrahelix,
\item describes a completely planar tetrahelix (the \emph{equitetrabeam}) useful as
  a structural beam or for applying standarod robot gaits to a tetrahedral robot, and
\item discusses the continuum of optimal tetrahelices between the BC helix and equitetrabeam.
\end{enumerate}
}

@@

\highlighttext{Engineers} would like to design nearly regular tetrahelices with a formula that
gives the vertices in space. Ideally, \highlighttext{this would be done}
by choosing the lengths of a small set of members.
In a space frame, this is a static design choice; in a Tetrobot, it is a
dynamic choice that can be used to twist the robot and/or exert linear or
angular force on the environment.

Ideally, \highlighttext{a simple formula would define} the vertices based on
any curvature or pitch we choose. It is a goal of
this paper to relate the Cartesian coordinate approach and the member-length approach to
generating a tetrahelix continuum.

@@

In order to develop the continuum of slightly irregular tetrahelices,
\highlighttext{a formula is needed}
 that gives us the vertices of just
one rail helix, denoted by color $c$ and integer vertex number $n$:
\[
(\forall n \in \mathbb{Z}, \forall c \in \{0,1,2\} : \vec{H}_{BCcolored}(n,c) = \vec{V}(3n +c)) .
\]

@@

In this formula, \highlighttext{integer} values of $n$ may be taken as a vertex number for one rail and used to compute
its Cartesian
coordinates. Allowing $n$ to take non-integer values defines a continuous
helix in space which is close to the segmented polyline of the outer
tetrahedra edges, and equals them at integer
values.

@@

\Cref{fig:helixcomparison} illustrates this difference with a 7-tetrahedra BC helix, which is
in fact the same geometry as the robot illustrated in \cref{fig:tetrobot}.
Although the vertices coincide,
 \cref{graycoxeter} evaluated
at real values generates the black helix which runs through every vertex, and \cref{eq:colored} defines
the red, yellow, and blue helices. In this figure,
these \highlighttext{continuous} rail helices have been rendered at a slightly higher radius than the vertices for clarity;
in actuality, the maximum distance between the continuous, curved helix and the
straight edges between vertices is much smaller than can be clearly rendered.

@@

Note, in \cref{railanglefig}, the $z$-axis travel for one rail edge is denoted by $d$. In \cref{graycoxeter} and  \cref{eq:colored}, the variable
$h$ is used for one third of the distance we name $d$. \highlighttext{The fact that $d = 3h$ is justified later}.
In this paper, we assume the length of a rail
is always $1$ as a simplification, except in proofs concerning rail length.
The rail length is a parameter in the JavaScript code in
\url{https://github.com/PubInv/tetrahelix/blob/master/js/tetrahelix_math.js} \cite{readtetrahelix}.

The $\vec{H}_{BCcolored}(n,c)$ formulation can be further clarified by rewriting directly in terms of the rail angle $\rho_{bc}$ rather than $\theta_{bc}$.
Intuitively, an expression where $c/3$ is multiplied by a $1/3$ rotation plus the rail angle $\rho$ \highlighttext{is sought}.
We expand 
the expressions $\theta_{bc}$ and $\rho_{bc}$ in \cref{eq:colored} and seek to isolate the term $c 2\pi/3 $.

@@

Thus, starting with the expression:
\begin{align*}
  c \theta_{bc} 
\end{align*}
introduce 3 into the denominator...
\begin{align*}
  (c/3)  (3 \theta_{bc}) 
\end{align*}
\highlighttext{seeking} $2\pi$ in the numerator, add canceling terms...
\begin{align*}
  (c/ 3) ((3 \theta_{bc} - 2 \pi)  + 2 \pi) 
\end{align*}
...and then use the definition of $\rho_{bc}$
\begin{align*}
  (c / 3) \rho_{bc}  + c 2 \pi /3 
\end{align*}
\highlighttext{to} finally obtain:
\begin{equation}
  c  ( \rho_{bc} +  2 \pi) /3  \text{.}
\end{equation}

This \highlighttext{allows \cref{eq:colored} to be rewritten}:

@@

Recall that $c \in \{0,1,2\}$, but $n$ is continuous (rational or real-valued).
\highlighttext{It can now be asserted that} in \cref{fig:helixcomparison} the black helix winds at
$\frac{3 \theta_{bc}}{\rho_{bc}} \approx 11.16 $ times the rate of a rail helix.

@@

The pitch is less than the number of tetrahedra because the tetrahedra
edges are not parallel to the axis of the tetrahelix.  It is a famous and interesting result
that the pitch is irrational. A BC helix never has two tetrahedra at
precisely the same orientation around the $z$-axis. However, this is
inconvenient to designers, who might prefer a rational pitch.
The idea of developing a rational period by arranging solid tetrahedra by relaxing
the face-to-face matching
has been explored\cite{sadler2013periodic}. 
\highlighttext{The next sections develop} slightly irregular edge lengths that support, for example,
a pitch of precisely 12
tetrahedra in one revolution which would allow an architect to
design a pleasing column having the top and bottom tetrahedra in the same relationship to the
viewer on both the capital and the basis.

@@

\highlighttext{
As engineers and roboticists it is desirable to design tetrahelices that
are slightly irregular that ``untwist'' the BC helix in an optimal way.
For machines
made out of linear actuators with limited range of extension,
the best definition of optimality is to minimize the difference in
lengths between the shortest and longest members in the robot of a given radius.
By maintaining all lengths as close to the same ``regular'' length as possible,
such as the mean length of the actuator,
one retains the greatest possible freedom of motion for the robot.
}

\highlighttext{
This paper use the term \emph{tetrahelix} to mean any structure physically constructible of
vertices and finite edges which is isomorphic to the BC helix, but with possibly different edge lengths,
and in which the
vertices lie on three helices (red, blue, and yellow, or ($R,B,Y$)). 
Mathematically, this structure extends
without bound, although physical robots or variable geometry trusses will
be truncations of such a structure.
This idea is captured by a function $T$ that gives the length between any
two nodes that are connected, and undefined ($\perp$) when not connected :
$T(n_1,n_2) : \mathbb{Z} \times \mathbb{Z} \mapsto \mathbb{R} \cup \{\perp\} $.
To describe a physically realizable machine, the function $T$ representing a tetrahelix
is constrained:
$(\forall n_1,n_2 : T(n_1,n_2) > 0)$.
For a BC helix, $(\forall x,y : T(x,y) \neq \perp \Rightarrow T(x,y) = 1)$.
}

\highlighttext{
The ratio of the longest member in a testrahelix to the shortest is:
\begin{equation}
R(T) = \frac{\max \{T(n_1,n_2) | n_1,n_2 \in \mathbb{Z} \}}{\min \{T(n_1,n_2) | n_1,n_2 \in \mathbb{Z} \}}
\label{eq:optdef}
\end{equation}
}

\highlighttext{
  The ratio of the BC helix is 1: $ R(H_{BC}) = 1$.
  }
\highlighttext{
A \emph{triple helix} is three congruent helices that share an axis. The remainder of this
section proves that
optimal tetrahelices are in fact triple helices.
In stages, this section proves that the helices in optimal tetrahelices:
\begin{enumerate}
\item have the same pitch,
\item have parallel axes,
\item share the same axis,
\item have the same radius,
  \item have the same rail length,
  \item have axially equidistant vertices, and therefore
  \item are in fact triple helices.
\end{enumerate}
The formula for radius that optimizes the tetrahelix for a given rail angle
is the result of the \cref{sec:parameterizing}.
}

\highlighttext{
\begin{theorem}
  All three helices in any physically realizable tetrahelix have the same pitch and parallel axes.
  \label{thm:pitchandparallel}
\end{theorem}
}

\begin{proof}
\highlighttext{  
Suppose that all three rails do not have the same pitch.
Let $n_1$ be a node on a  shortest edge between
two rails. Let $n_2$ be a node on a different rail.
Increasing the difference $\|n_2-n_1\|$ increases edge lengths $T(n_1,n_2)$
without bound,
which is not physically realizable.
}

\highlighttext{  
  Similarly, if the axes are not parallel, for any length $\lambda$ there is an edge of $T(x,y)$
  such that $T(x,y) > \lambda$, which is not physical.
}
\end{proof}

\highlighttext{
\Cref{thm:pitchandparallel} justifies talking about the singular
\emph{pitch} of 
the optimal tetrahelix as the pitch of its three rail helices $(R,B,Y)$, even though there are
three such helices of equivalent pitch.
Since the axes of helices of any physical tetrahelix are parallel, take
them to be parallel to the $z$ axis. Define the radius of the tetrahelix:
\begin{multline*}
rad(T) = \text{the radius of the smallest cylinder} \\
\text{which contains all of nodes of the tetrahelix } T
\end{multline*}
Define a \emph{minimax edge-length optimal tetrahelix} or just an
\emph{optimal tetrahelix} to be a tetrahelix for which there exists
no other tetrahelix  of the same radius with lower ratio of longest edge length to
shortest edge length
\begin{equation}
  OPT(T) \equiv \neg (\exists S : R(S) < R(T) \wedge rad(S) = rad(T))
  \label{eq:optdef}
\end{equation}
}

\highlighttext{
\begin{lemma}
  A tetrahelix where all points lie on the same line is not optimal.
    \label{lemma:notpoint}
\end{lemma}
\begin{proof}
As a little lemma for the proof below, observe that a tetrahelix $T$ of zero radius, where
all points lie on the same line,
is not as optimal as a tetrahelix of a small radius. The edges between rails $(R,B,Y)$ will be
shorter than the rail edges, and moving them apart slightly to form $T'$ lengthens the between-edge
rails, improving the ratios, so $R(T') < R(T)$.
\end{proof}
}

In the proof, below, \highlighttext{it is} useful to consider projection diagrams that are the axial projection
of a tetrahelix onto the $XY$-plane. \Cref{fig:projectiondiagram,fig:wobble} are examples
of such a diagram.
\begin{lemma}
  If the rail angle $0 < \rho < \pi$ is a rational multiple of $\pi$, then the projection of
  edges in a helix of that rail angle  along the $z$-axis onto the $XY$-plane form a
  regular polygon of 3 or more sides, \highlighttext{otherwise} they fill in a complete circle.
  \label{lemma:ngon}
\end{lemma}
\begin{proof}
  All points lying on a helix projected along the axis lie on a circle in the $XY$-plane.
\highlighttext{(By \cref{lemma:notpoint}, the circle is not a simple point.)}
  Helices are periodic in the $z$ dimension modulo $2\pi$.
  If $2\pi/\rho$ is irrational, the projection onto the $XY$-plane will contain
  an unbounded number of points on a circle.
  If and only if $2\pi/\rho$ is rational, the projection onto
  the $XY$-plane will contain a finite number of points.
  Because $\pi$ is transcendental and irrational,
  $2\pi/\rho$ is rational if and only if $\rho = a\pi/b$, where $a$ and $b$ are integers
  and, \highlighttext{assume} without loss of generality, $a$ and $b$ are coprime.
  Since $\rho < \pi$, therefore $a < b$. Also, $\rho > 0$, therefore $a > 0$.
  The number of points in the projection is $2b$ if $a$ is odd, and $b$ if $a$ is even.
  This polygon has at least 3 sides, since either $\rho$ is irrational or $b > a$, and therefore
  $b \geq 2$. If $a/b = 1/2$, the projection is a square, which has four sides. 
\end{proof}

\begin{theorem}
  Any optimal tetrahelix with a rail angle of magnitude less than $\pi$ has all three
  axes coincident.
  \label{thm:coincident}
\end{theorem}
\begin{proof}
  Case 1: Suppose that $\rho$ is zero.
  Each helix has zero curvature, that is, it is a straight line. These lines are eqivalent
  to some three degenerate helices with coincident axes, possibly with different radii,
  so long as there is a phase
  term in the defintion of the helix, as in \cref{eq:colored}. \highlighttext{That the
  radii must be eqivalent is shown later}.

@@

  \highlighttext{By symmetry} without loss of generality, define the Red helix to have its axis on the $z$-axis.
  Since there must be at least one Red-to-Yellow or a Red-to-Blue edge that is either a minimum or a maximum,
  without loss of generality, define the Blue helix to be a helix that
  has an edge connection to the Red helix that is either a maximum or a minimum.
  Let $B'$ be a translation in the $XY$-plane of the blue helix $B$  so that its axis is the $z$-axis and
  coincident with the red helix $R$. Let $D$ be the distance between the axis of the Blue helix
  $B$ and $B'$. We will show that if $D>0$ then $B$ ``wobbles'' in a way that cannot be optimal.
  Define a wobble vector by:
  \begin{equation}
    \vec{W}(n) = \vec{B}(n) - \vec{B}'(n) \text{ .}
  \end{equation}
  where $\vec{B}(n)$ and $\vec{B'}(n)$ is the vector
  $\begin{bmatrix}
    x \\
    y 
  \end{bmatrix}$
  for the projection of the $n$th vertex of $B$ and $B'$. 
  Note that $|| \vec{R}(n) - \vec{B'}(n+k)||$ (the Euclidean distance of the vertices)
  is a constant for any $k$, because $R$ and $B'$ have the
  same pitch and the same axis, even if they do not have the same radius.

@@

Therefore, a tetrahelix with different radii is not as optimal as some two-class 
tetrahelix generated by \cref{eq:triple}, \highlighttext{ and so by \cref{eq:optdef}} is not optimal.

>>>>

% \documentclass[review]{siamonline1116}
\documentclass[10pt,final]{journals-1.0/asme2ej}
% \usepackage{geometry} % see geometry.pdf on how to lay out the page. There's lots.
% \usepackage{hyperref}
\usepackage{graphicx}
\usepackage{gensymb}
% \usepackage[affil-it]{authblk}
% \usepackage[toc,page]{appendix}
% \usepackage{pifont}
\usepackage{amsmath}
% \usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{hyperref}
\usepackage{cleveref}

@@

@@

\title{Transforming Optimal Tetrahelices between the Boerdijk-Coxeter Helix and a Planar-faced Tetrahelix}
\author{Robert L. Read, PhD
  \affiliation{
    Founder, Public Invention \\
    Austin, TX, 78704 \\
    Email: \href{mailto:read.robert@gmail.com}{read.robert@gmail.com} 
    }
}

@@

%% TODO: Add something about Tetrobots into abstract.
\begin{abstract}
  The Boerdijk-Coxeter helix (BC helix, or tetrahelix) is a
  face-to-face stack of regular tetrahedra forming a helical column.  Treating the edges of
  these tetrahedra as structural members creates an attractive and
  inherently rigid space frame, and therefore is interesting to architects,
  mechanical engineers, and roboticists.  A formula is developed that matches the
  visually apparent helices forming the outer rails of the BC helix.
  This formula is generalized to a formula convenient to designers.
  Formulae for 
  computing the
  parameters that give proven edge-length minimax-optimal tetrahelices
  are given,
  allowing transformation through a continuum of 
  of optimum tetrahelices of varying curvature while maximizing regularity.
  The endpoints of this continuum are the BC helix and
  a structure of zero curvature, the \emph{equitetrabeam}.
 Only one out of three members in the system change their length to transform
  the structure into
  any point in the continuum.
  Numerically finding the rail angle from the equation for
  pitch allows optimal tetrahelices of any pitch to be designed. 
  An interactive tool for such design and experimentation is provided: \url{https://pubinv.github.io/tetrahelix/}.
  A formula for the inradius of optimal tetrahelices is given.
  The continuum allows a regular Tetrobot supporting a length change of less than $16\%$ in the
  BC configuration to untwist into a hexapodal or $n$-podal robot
  to use standard gaits.
\end{abstract}

@@

The Boerdijk-Coxeter helix\cite{coxeter1985simplicial} (BC helix) (see \cref{fig:closeup,fig:closeuporthog}), is
a face-to-face stack of tetrahedra that winds about a straight axis.
Architects, structural engineers, and roboticists are inspired
by and follow such regular mathematical models.
However, since they can also build structures and
machines from members of differing or even dynamically changing length, it is
useful to develop the mathematics of structures formed from tetrahedra
where one relaxes regularity.

@@

Buckminster Fuller called the BC helix a \emph{tetrahelix}\cite{fuller1982synergetics},
a term now commonly used. This paper reserves \emph{BC helix} to mean the purely
regular structure and use \emph{tetrahelix} to refer
to any structure isomorphic to the BC helix.

@@

A BC helix does not rest stably on a plane. It is convenient to
be able to ``untwist'' it and to form a tetrahelix space frame that has a
flat planar surface. By making length changes in a certain way, a tetrahelix untwists to form a \emph{tetrabeam} which has planar faces
and has, for example, an equilateral triangular profile. This paper
develops the equations needed to untwist the tetrahelix. All math
developed here is available in JavaScript and demonstrated by an interactive
design website \url{https://pubinv.github.io/tetrahelix/}\cite{readtetrahelix}
and a video\cite{tetrahelixvideo},
from which \cref{fig:closeup,fig:closeuporthog} and later figures in this paper 
are taken.

@@

\Cref{fig:series} displays a continuum of tetrahelices optimal in the sense of being
as regular as possible,
which is the main result of this
paper. The closest helix is the BC helix, and the furthest
is the equitetrabeam, defined in \cref{sec:equitetrabeam} and \cref{fig:naming,fig:equitetrabeam}.

@@

However, a robot must move, and so it is interesting to consider the
transmutations of these regular geometries. Developing a functional
gait for our physical Tetrobot (the 7-Tet Glussbot) was in fact the motivation
for this work.
By changing only the length of only one third of the members,
one can dynamically untwist a physical Tetrobot
forming the Boerdijk-Coxeter configuration into the equitetrabeam which rests flat on the plane,
as shown in figs. 4 and 5.

By untwisting the tetrahelix so that is has a planar surface resting on the ground,
each vertex touching the ground becomes a foot or pseudopod. A robot can thus
become a hexapod or $n$-pod Tetrobot, and the already well-developed approaches to
hexapod gaits may be applied to make the robot walk or crawl.

@@

The remainder of this  paper:
\begin{enumerate}
\item provides a new formulation for the BC helix that is more natural for engineers,
\item defines a concept of minimum maximum edge-length optimality that maintains
  the structure as close to regularity as possible,
  \item proves that all
    optimal tetrahelices are in fact triple helices,
\item provides formulae for constructing optimal tetrahelices of different twists or for optimally
    transforming a
    robot from a twisted to a planar configuration in a way that maintains optimum regularity,
\item gives a formula for the inradius of an optimal tetrahelix,
\item describes a completely planar tetrahelix (the \emph{equitetrabeam}) useful as
  a structural beam or for applying standarod robot gaits to a tetrahedral robot, and
\item discusses the continuum of optimal tetrahelices between the BC helix and equitetrabeam.
\end{enumerate}

@@

Engineers would like to design nearly regular tetrahelices with a formula that
gives the vertices in space. Ideally, this would be done
by choosing the lengths of a small set of members.
In a space frame, this is a static design choice; in a Tetrobot, it is a
dynamic choice that can be used to twist the robot and/or exert linear or
angular force on the environment.

Ideally, a simple formula would define the vertices based on
any curvature or pitch we choose. It is a goal of
this paper to relate the Cartesian coordinate approach and the member-length approach to
generating a tetrahelix continuum.

@@

In order to develop the continuum of slightly irregular tetrahelices,
a formula is needed
that gives us the vertices of just
one rail helix, denoted by color $c$ and integer vertex number $n$:
\[
(\forall n \in \mathbb{Z}, \forall c \in \{0,1,2\} : \vec{H}_{BCcolored}(n,c) = \vec{V}(3n +c)) .
\]

@@

In this formula, integer values of $n$ may be taken as a vertex number for one rail and used to compute
its Cartesian
coordinates. Allowing $n$ to take non-integer values defines a continuous
helix in space which is close to the segmented polyline of the outer
tetrahedra edges, and equals them at integer
values.

@@

\Cref{fig:helixcomparison} illustrates this difference with a 7-tetrahedra BC helix, which is
in fact the same geometry as the robot illustrated in \cref{fig:tetrobot}.
Although the vertices coincide,
 \cref{graycoxeter} evaluated
at real values generates the black helix which runs through every vertex, and \cref{eq:colored} defines
the red, yellow, and blue helices. In this figure,
these continuous rail helices have been rendered at a slightly higher radius than the vertices for clarity;
in actuality, the maximum distance between the continuous, curved helix and the
straight edges between vertices is much smaller than can be clearly rendered.

@@

Note, in \cref{railanglefig}, the $z$-axis travel for one rail edge is denoted by $d$. In \cref{graycoxeter} and  \cref{eq:colored}, the variable
$h$ is used for one third of the distance we name $d$. The fact that $d = 3h$ is justified later.
In this paper, we assume the length of a rail
is always $1$ as a simplification, except in proofs concerning rail length.
The rail length is a parameter in the JavaScript code in
\url{https://github.com/PubInv/tetrahelix/blob/master/js/tetrahelix_math.js} \cite{readtetrahelix}.

The $\vec{H}_{BCcolored}(n,c)$ formulation can be further clarified by rewriting directly in terms of the rail angle $\rho_{bc}$ rather than $\theta_{bc}$.
Intuitively, an expression where $c/3$ is multiplied by a $1/3$ rotation plus the rail angle $\rho$ is sought.
We expand 
the expressions $\theta_{bc}$ and $\rho_{bc}$ in \cref{eq:colored} and seek to isolate the term $c 2\pi/3 $.

@@

Thus, starting with the expression:
\begin{align*}
  c \theta_{bc} 
\end{align*}
introduce 3 into the denominator...
\begin{align*}
  (c/3)  (3 \theta_{bc}) 
\end{align*}
seeking $2\pi$ in the numerator, add canceling terms...
\begin{align*}
  (c/ 3) ((3 \theta_{bc} - 2 \pi)  + 2 \pi) 
\end{align*}
...and then use the definition of $\rho_{bc}$
\begin{align*}
  (c / 3) \rho_{bc}  + c 2 \pi /3 
\end{align*}
to finally obtain:
\begin{equation}
  c  ( \rho_{bc} +  2 \pi) /3  \text{.}
\end{equation}

This allows \cref{eq:colored} to be rewritten:

@@

Recall that $c \in \{0,1,2\}$, but $n$ is continuous (rational or real-valued).
It can now be asserted that in \cref{fig:helixcomparison} the black helix winds at
$\frac{3 \theta_{bc}}{\rho_{bc}} \approx 11.16 $ times the rate of a rail helix.

@@

The pitch is less than the number of tetrahedra because the tetrahedra
edges are not parallel to the axis of the tetrahelix.  It is a famous and interesting result
that the pitch is irrational. A BC helix never has two tetrahedra at
precisely the same orientation around the $z$-axis. However, this is
inconvenient to designers, who might prefer a rational pitch.
The idea of developing a rational period by arranging solid tetrahedra by relaxing
the face-to-face matching
has been explored\cite{sadler2013periodic}. 
The next sections develop slightly irregular edge lengths that support, for example,
a pitch of precisely 12
tetrahedra in one revolution which would allow an architect to
design a pleasing column having the top and bottom tetrahedra in the same relationship to the
viewer on both the capital and the basis.

@@

As engineers and roboticists it is desirable to design tetrahelices that
are slightly irregular that ``untwist'' the BC helix in an optimal way.
For machines
made out of linear actuators with limited range of extension,
the best definition of optimality is to minimize the difference in
lengths between the shortest and longest members in the robot of a given radius.
By maintaining all lengths as close to the same ``regular'' length as possible,
such as the mean length of the actuator,
one retains the greatest possible freedom of motion for the robot.

This paper use the term \emph{tetrahelix} to mean any structure physically constructible of
vertices and finite edges which is isomorphic to the BC helix, but with possibly different edge lengths,
and in which the
vertices lie on three helices (red, blue, and yellow, or ($R,B,Y$)). 
Mathematically, this structure extends
without bound, although physical robots or variable geometry trusses will
be truncations of such a structure.
This idea is captured by a function $T$ that gives the length between any
two nodes that are connected, and undefined ($\perp$) when not connected :
$T(n_1,n_2) : \mathbb{Z} \times \mathbb{Z} \mapsto \mathbb{R} \cup \{\perp\} $.
To describe a physically realizable machine, the function $T$ representing a tetrahelix
is constrained:
$(\forall n_1,n_2 : T(n_1,n_2) > 0)$.
For a BC helix, $(\forall x,y : T(x,y) \neq \perp \Rightarrow T(x,y) = 1)$.

The ratio of the longest member in a testrahelix to the shortest is:
\begin{equation}
R(T) = \frac{\max \{T(n_1,n_2) | n_1,n_2 \in \mathbb{Z} \}}{\min \{T(n_1,n_2) | n_1,n_2 \in \mathbb{Z} \}}
\label{eq:optdef}
\end{equation}

 The ratio of the BC helix is 1: $ R(H_{BC}) = 1$.

A \emph{triple helix} is three congruent helices that share an axis. The remainder of this
section proves that
optimal tetrahelices are in fact triple helices.
In stages, this section proves that the helices in optimal tetrahelices:
\begin{enumerate}
\item have the same pitch,
\item have parallel axes,
\item share the same axis,
\item have the same radius,
  \item have the same rail length,
  \item have axially equidistant vertices, and therefore
  \item are in fact triple helices.
\end{enumerate}
The formula for radius that optimizes the tetrahelix for a given rail angle
is the result of the \cref{sec:parameterizing}.

\begin{theorem}
  All three helices in any physically realizable tetrahelix have the same pitch and parallel axes.
  \label{thm:pitchandparallel}
\end{theorem}

\begin{proof}
Suppose that all three rails do not have the same pitch.
Let $n_1$ be a node on a  shortest edge between
two rails. Let $n_2$ be a node on a different rail.
Increasing the difference $\|n_2-n_1\|$ increases edge lengths $T(n_1,n_2)$
without bound,
which is not physically realizable.

Similarly, if the axes are not parallel, for any length $\lambda$ there is an edge of $T(x,y)$
such that $T(x,y) > \lambda$, which is not physical.

\end{proof}

\Cref{thm:pitchandparallel} justifies talking about the singular
\emph{pitch} of 
the optimal tetrahelix as the pitch of its three rail helices $(R,B,Y)$, even though there are
three such helices of equivalent pitch.
Since the axes of helices of any physical tetrahelix are parallel, take
them to be parallel to the $z$ axis. Define the radius of the tetrahelix:
\begin{multline*}
rad(T) = \text{the radius of the smallest cylinder} \\
\text{which contains all of nodes of the tetrahelix } T
\end{multline*}
Define a \emph{minimax edge-length optimal tetrahelix} or just an
\emph{optimal tetrahelix} to be a tetrahelix for which there exists
no other tetrahelix  of the same radius with lower ratio of longest edge length to
shortest edge length
\begin{equation}
  OPT(T) \equiv \neg (\exists S : R(S) < R(T) \wedge rad(S) = rad(T))
  \label{eq:optdef}
\end{equation}

\begin{lemma}
  A tetrahelix where all points lie on the same line is not optimal.
    \label{lemma:notpoint}
\end{lemma}
\begin{proof}
As a little lemma for the proof below, observe that a tetrahelix $T$ of zero radius, where
all points lie on the same line,
is not as optimal as a tetrahelix of a small radius. The edges between rails $(R,B,Y)$ will be
shorter than the rail edges, and moving them apart slightly to form $T'$ lengthens the between-edge
rails, improving the ratios, so $R(T') < R(T)$.
\end{proof}

In the proof, below, it is useful to consider projection diagrams that are the axial projection
of a tetrahelix onto the $XY$-plane. \Cref{fig:projectiondiagram,fig:wobble} are examples
of such a diagram.
\begin{lemma}
  If the rail angle $0 < \rho < \pi$ is a rational multiple of $\pi$, then the projection of
  edges in a helix of that rail angle  along the $z$-axis onto the $XY$-plane form a
  regular polygon of 3 or more sides, otherwise they fill in a complete circle.
  \label{lemma:ngon}
\end{lemma}
\begin{proof}
  All points lying on a helix projected along the axis lie on a circle in the $XY$-plane.
(By \cref{lemma:notpoint}, the circle is not a simple point.)
  Helices are periodic in the $z$ dimension modulo $2\pi$.
  If $2\pi/\rho$ is irrational, the projection onto the $XY$-plane will contain
  an unbounded number of points on a circle.
  If and only if $2\pi/\rho$ is rational, the projection onto
  the $XY$-plane will contain a finite number of points.
  Because $\pi$ is transcendental and irrational,
  $2\pi/\rho$ is rational if and only if $\rho = a\pi/b$, where $a$ and $b$ are integers
  and, assume without loss of generality, $a$ and $b$ are coprime.
  Since $\rho < \pi$, therefore $a < b$. Also, $\rho > 0$, therefore $a > 0$.
  The number of points in the projection is $2b$ if $a$ is odd, and $b$ if $a$ is even.
  This polygon has at least 3 sides, since either $\rho$ is irrational or $b > a$, and therefore
  $b \geq 2$. If $a/b = 1/2$, the projection is a square, which has four sides. 
\end{proof}

\begin{theorem}
  Any optimal tetrahelix with a rail angle of magnitude less than $\pi$ has all three
  axes coincident.
  \label{thm:coincident}
\end{theorem}
\begin{proof}
  Case 1: Suppose that $\rho$ is zero.
  Each helix has zero curvature, that is, it is a straight line. These lines are eqivalent
  to some three degenerate helices with coincident axes, possibly with different radii,
  so long as there is a phase
  term in the defintion of the helix, as in \cref{eq:colored}. That the
  radii must be eqivalent is shown later.

@@

  By symmetry without loss of generality, define the Red helix to have its axis on the $z$-axis.
  Since there must be at least one Red-to-Yellow or a Red-to-Blue edge that is either a minimum or a maximum,
  without loss of generality, define the Blue helix to be a helix that
  has an edge connection to the Red helix that is either a maximum or a minimum.
  Let $B'$ be a translation in the $XY$-plane of the blue helix $B$  so that its axis is the $z$-axis and
  coincident with the red helix $R$. Let $D$ be the distance between the axis of the Blue helix
  $B$ and $B'$. We will show that if $D>0$ then $B$ ``wobbles'' in a way that cannot be optimal.
  Define a wobble vector by:
  \begin{equation}
    \vec{W}(n) = \vec{B}(n) - \vec{B}'(n) \text{ .}
  \end{equation}
  where $\vec{B}(n)$ and $\vec{B'}(n)$ is the vector
  $\begin{bmatrix}
    x \\
    y 
  \end{bmatrix}$
  for the projection of the $n$th vertex of $B$ and $B'$. 
  Note that $|| \vec{R}(n) - \vec{B'}(n+k)||$ (the Euclidean distance of the vertices)
  is a constant for any $k$, because $R$ and $B'$ have the
  same pitch and the same axis, even if they do not have the same radius.

@@

Therefore, a tetrahelix with different radii is not as optimal as some two-class 
tetrahelix generated by \cref{eq:triple}, and so by \cref{eq:optdef} is not optimal.
